17th International Satisfiability Modulo Theories Competition (SMT-COMP 2022): Rules and Procedures

### 2022-02-08 rules/rules22.tex
more
[rules] adding changes for this year

\newcommand{\maintrack}{Single Query Track\xspace}
\newcommand{\inctrack}{Incremental Track\xspace}
\newcommand{\ucoretrack}{Unsat-Core Track\xspace}
\newcommand{\mvaltrack}{Model-Validation Track\xspace}
\newcommand{\challtrack}{Industry-Challenge Track\xspace}
\newcommand{\paralleltrack}{Parallel Track\xspace}
\newcommand{\cloudtrack}{Cloud Track\xspace}

@@

SMT-COMP~2022 is part of the SMT Workshop~2022
(\url{http://smt-workshop.cs.uiowa.edu/2022/}) \hb{check},
which is affiliated with IJCAR~2022 (\url{https://easychair.org/smart-program/IJCAR2022/}).
The SMT Workshop will include a block of time to present the results of the
competition.
%
Accordingly, researchers are highly encouraged to submit both new
benchmarks and new or improved solvers to raise the level of
competition and advance the state of the art in automated SMT problem
solving.

SMT-COMP 2022 will have five \hb{check} tracks: the \maintrack (before 2019: Main Track),
the \inctrack (before 2019: Application Track), the \ucoretrack, and the
\mvaltrack.  In addition there will be two experimental tracks sponsored
by AWS: the \paralleltrack{} and the \cloudtrack{}.
%
Within each track there are multiple divisions, where each division
uses benchmarks from a specific group of SMT-LIB logics.
We will recognize winners as measured by number of benchmarks solved
(taking into account the weighting detailed in
Section~\ref{sec:scoring}); we will also recognize solvers based on
additional criteria.

@@

\hb{change more from here}
As in previous years, we have revised the rules slightly.  Some rule
changes in 2021 are designed to not overly punish solvers producing
syntactically wrong output, by treating those outputs as unknown and
not as unsoundness.  Moreover, we added more divisions to the model
validation track.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf Divisions now may contain more than one logic.} Previously, every
  division was comprised of a single logic. There are currently over 60 logics
  in SMT-LIB, which makes a short summary of the results impossible.  On the
  other hand the competition-wide scoring does a comparison between solvers that
  cannot be sensibly compared as they support completely disjoint logics. To get
  a better overview, this year divisions will combine related logics, thus
  reducing the number of divisions and obtaining more manageable and meaningful
  results.  Rather than logics, solvers will declare the divisions they
  enter. They will be ranked by the number of benchmarks they solve across all
  logics in the divisions entered.
  \begin{itemize}
    \item Quantifier-free divisions:
    \begin{itemize}
      \item QF\_Equality\\
Logics: QF\_UF, QF\_AX, QF\_DT, QF\_UFDT
\item QF\_Equality+LinearArith\\
Logics: QF\_ALIA, QF\_AUFLIA, QF\_UFLIA, QF\_UFLRA, QF\_UFIDL
\item QF\_Equality+NonLinearArith\\
Logics: QF\_UFNRA, QF\_UFNIA, QF\_ANIA, QF\_AUFNIA
\item QF\_Equality+Bitvec\\
Logics: QF\_ABV, QF\_UFBV, QF\_AUFBV
\item QF\_Equality+Bitvec+Arith\\
Logics: QF\_AUFBVLIA, QF\_AUFBVNIA, QF\_UFBVLIA
\item QF\_LinearIntArith\\
Logics: QF\_LIA, QF\_LIRA, QF\_IDL
\item QF\_LinearRealArith\\
Logics: QF\_LRA, QF\_RDL
\item QF\_Bitvec\\
Logics: QF\_BV
\item QF\_FPArith\\
Logics: QF\_FP, QF\_UFFP, QF\_FPLRA, QF\_BVFP, QF\_ABVFP, QF\_AUFBVFP, QF\_BVFPLRA, QF\_ABVFPLRA
\item QF\_NonLinearIntArith\\
Logics: QF\_NIA, QF\_NIRA
\item QF\_NonLinearRealArith\\
Logics: QF\_NRA
\item QF\_Strings\\
Logics: QF\_S, QF\_SLIA, QF\_SNIA

    \end{itemize}
    \item Divisions with quantifiers:
    \begin{itemize}
      \item Equality\\
Logics: UF, UFDT
\item Equality+LinearArith\\
Logics: \sloppy ALIA, AUFLIA, UFLIA, UFIDL, AUFLIRA, UFLRA, UFDTLIA, UFDTLIRA, AUFDTLIA, AUFDTLIRA
\item Equality+MachineArith\\
Logics: AUFFPDTLIRA, UFFPDTLIRA, UFFPDTNIRA, ABVFP, ABVFPLRA, AUFBV, AUFBVFP,
AUFBVDTLIA, UFBV, UFBVFP, UFBVLIA
\item Equality+NonLinearArith\\
Logics: ANIA, AUFDTNIRA, UFDTNIA, UFDTNIRA, AUFNIA, AUFNIRA, UFNIA, UFNRA
\item Arith\\
Logics: LRA, LIA, NIA, NRA
\item Bitvec\\
Logics: BV
\item FPArith\\
Logics: BVFP, FP, BVFPLRA, FPLRA
    \end{itemize}
  \end{itemize}

  \rationale{We are aware that any partitioning of logics into divisions can be
    disadvantageous to some solvers that support some but not all logics in the
    division. We believe however that to a lesser extent this situation is already
    common, thinking of different features among problems in the same logic. So in
    our view the disadvantages are offset by the benefit of having a better
    presentation of the results. Regardless, results per logic will still be
    accessible within the results page of a given division, similarly to how one can
    now see the different scores other than sequential for a given logic.

    The divisions above were defined according to the following rationale:

    \begin{itemize}
      \item Similar logics that do not standout by themselves were combined.
      \item Integer and real, as well as bit-vector and floating-point,
      arithmetics are separated
      \begin{itemize}
        \item Note that the ``Arith'' division goes against this. However, its
        composing logics do not standout out by themselves as currently
        represented in SMT-LIB.
      \end{itemize}
      \item Arrays and datatypes are handled in a similar enough way to UF to be combined with it.
        \item Floating-point logics are combined since generally this is the dominating component for any logic contaning FP.
    \end{itemize}
  }

  \item {\bf The option \akey{:print-success} is explicitly set to
    true or false.}  Previous competitions set this option to
    \akey{true} for the incremental track.  In addition, this
    competition will set the option to \akey{false} for all other
    tracks.
    \rationale{According to the standard the default value of
      \akey{:print-success} is true.  So a conforming solver would
      unnecessarily print \akey{success} after each command that needs
      to be stored and skipped by the post-processors.  Additionally,
      solvers can use this option as indication whether incremental
      checking is desired.}

    We were asked to set the \akey{:incremental} option to true for
    the incremental track.  However this is problematic, because it is
    not a standard option and solvers may not support it.

  \item {\bf Error result is for unsoundness, only.}
    Previous competitions marked malformed output in the unsat core and
    the model validation track as errors.  These errors basically disqualify
    a solver that outputs a slightly malformed model.  This year the error
    result is only given for unsat cores that are not unsatisfiable or models
    that don't satisfy the formula. Of courses, answering sat in unsat
    core track or unsat in model validation track is still considered an
    error.
    \rationale{An error result is very punishing and should be reserved
      for problems that cannot be easily detected.  Giving a syntactically
      wrong output, e.g. an error message, should just be considered as
      not solving the task, not as providing an unsound result.}

  \item {\bf The organizers may manually check unsound results.}
    If solvers disagree on a result, the organizers reserve the right
    to play the judge and determine which solver is unsound.  The unsound
    solver is punished, the sound solver gets the point.
    Previous competitions would take benchmark out of the scoring.
    \rationale{Often, it can be decided which solver is unsound, even when
      the benchmark has no status.  For example, when a solver was unsound in
      other divisions and the authors submitted a fixed version, or because
      a solver returning sat can be asked for a model.  In that case it would
      be bad if solver that is known to be unsound
      won the competition, just because the benchmarks it was unsound
      on had no status.}

  \item {\bf Punishment of unsound results is adapted to fix sat/unsat scores.}

    Previously an unsound response for a benchmark did punish the
    score corresponding to the benchmark status.  We fix it to count
    errors to the score that correspond to the solver output, not the
    benchmark status.
    \rationale{If a solver returns \texttt{unsat} for a
      satisfiable benchmark, all other \texttt{unsat} responses of the
      solver are no longer trustworthy.  So it is the unsat score of
      the solver that must be punished.  The \texttt{sat} responses
      may still be trustworthy.  In the extreme case, a solver that
      always return \texttt{unsat} would a high error score for the
      unsat category under the new rules, and would get no points at
      all for the sat category.}

  \item {\bf Benchmark Selection.}
      Apart from the \cloudtrack{} and the \paralleltrack{}, this year we
        increase the percentage of the benchmarks that are run in the
        competition back to 50 \%.
    \rationale{We feel that increasing the number of benchmarks is
    manageable and 50 \% is a more round number than 40 \%.  The
    \cloudtrack{} and \paralleltrack{} need to be run on a more
    restricted set because of the amount of parallel resources we
    provide for the tracks.}

  \item {\bf Stricter checks for output in single query track.}
    It is now checked that the output in the single query track starts with
    sat or unsat.  Error messages before such an output is no longer ignored
    and are treated as an \texttt{unknown} result.
    {\bf Note:} Solver authors should make sure not to print any debugging
    output.  This includes debugging output on stderr, because the
    post-processors cannot distinguish between stderr and stdout.

  \rationale{A solver that does not support the full logic may
    report errors and then continue ignoring the failed assertion.  This
    way, it may be flagged as unsound under the old rules, even though it
    clearly indicated the error.}

  \item {\bf \mvaltrack.}  \emergencystretch.5cm Last year's competition introduced the experimental
  divisions QF\_IDL, QF\_RDL, QF\_LIA, QF\_LRA, QF\_LIRA in the model validation
  track.  This year, the divisions containing these logics will no longer be
  experimental.  In addition, we will add this year new experimental divisions
  QF\_Equality (only with QF\_UF benchmarks), QF\_Equality+LinearArith (only with
  QF\_UFIDL, QF\_UFLIA, and QF\_UFLRA benchmarks), and QF\_Equality+Bitvec (only with
  QF\_UFBV benchmarks).

  \rationale{There were only small issues with last year's divisions, so there
    is no reason to keep them experimental. For the new divisions, as before,
    given the inconsistencies across different model producing solvers, we
    proceed in an experimental fashion to push for model standardization. Since
    both arrays and datatypes add further complications on top of uninterpreted
    functions, we exclude benchmarks containing those.}

  \item {\bf Model syntax in \mvaltrack.}  The syntax used in the
    model validation track did not conform to the SMT-LIB standard.
    The difference is that the syntax used previously included an
    additional keyword \texttt{model} after the opening parenthesis in
    the output.  This year we support both the SMT-LIB syntax and the
    old syntax.  We urge solver writers to upgrade their solver to use
    the new syntax.

    We added new tracks for UF logics. According to the SMT-LIB standard,
    model values of uninterpreted sorts should use the syntax
    \texttt{(as @name Sort)}.
    In particular the model validator needs to know the sort of the model
    value.  Furthermore, it is implicitly assumed that model values
    differ if they have different names.

@@

@@

\header{Disclosure.}
%
Haniel Barbosa is part of the developing teams of the SMT solvers
cvc5~\cite{cvc4-smtcomp18} and veriT~\cite{verit}.
%
Fran\c{c}ois is part of the developing team of the SMT solver COLIBRI~\cite{colibri}.
%
Jochen Hoenicke is part of the developing team of the SMT solver
SMTInterpol~\cite{smtinterpol}.

>>>>

\newcommand{\maintrack}{Single Query Track\xspace}
\newcommand{\inctrack}{Incremental Track\xspace}
\newcommand{\ucoretrack}{Unsat-Core Track\xspace}
\newcommand{\mvaltrack}{Model-Validation Track\xspace}
\newcommand{\challtrack}{Industry-Challenge Track\xspace}
\newcommand{\paralleltrack}{Parallel Track\xspace}
\newcommand{\cloudtrack}{Cloud Track\xspace}
\newcommand{\prooftrack}{Proof Exhibition Track\xspace}

@@

SMT-COMP~2022 is part of the SMT Workshop~2022
(\url{http://smt-workshop.cs.uiowa.edu/2022/}),
which is affiliated with IJCAR~2022 (\url{https://easychair.org/smart-program/IJCAR2022/}).
The SMT Workshop will include a block of time to present the results of the
competition.
%
Accordingly, researchers are highly encouraged to submit both new
benchmarks and new or improved solvers to raise the level of
competition and advance the state of the art in automated SMT problem
solving.

SMT-COMP 2022 will have seven tracks: the \maintrack (before 2019: Main Track),
the \inctrack (before 2019: Application Track), the \ucoretrack, the \mvaltrack,
the \paralleltrack{} and the \cloudtrack{} (both sponsered by Amazon Web
Services), and an experimental \prooftrack{} focused on a qualitative evaluation
of different proof formats, proof checkers and proof-producing solvers.
%
Within each track there are multiple divisions, where each division
uses benchmarks from a specific group of SMT-LIB logics.
%
We will recognize winners in all tracks but the \prooftrack{}.
%
They will be determined by the number of benchmarks solved (taking into account
the weighting detailed in Section~\ref{sec:scoring}); we will also recognize
solvers based on additional criteria.

@@

As in previous years, we have revised the rules slightly.  Some rule
changes in 2022 are designed to improve how logics are organized into
divisions. We have also added more divisions to the \mvaltrack and have defined a
new \prooftrack.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf Changes in divisions.} Last year we changed
  divisions to be composed of one or more related logics.
  %
  This year we will keep this organization but we will separate logics into
  divisions slightly differently: we will have a QF\_Datatypes division with the
  logics QF\_DT, QF\_UFDT and the QF\_Equality division now only contains the
  QF\_UF and QF\_AX logics.

  \rationale{The feedback from the participants in last year's competition was
    that datatypes were too dissimilar from QF\_UF and QF\_AX to be organized in
    the same division.}

  \item {\bf Logics can be non-competitive in competitive divisions.} Last year
  we considered every logic within a division to competitive unless the division
  as whole was non-competitive. This year if a logic is non-competitive within a
  division (i.e., only one competitive solver enrolled in it), it will not be
  ran.

  \rationale{We believe the previous setup could benefit too much a solver that
    supports more logics that its competitor.
    %
    In particular there was a whose winner was defined by the benchmarks of a
    non-competitive logic, which we believe is not ideal.
    %
    Note that the incentive for solvers to support more logics continues since
    as soon as there is more than one solver supporting a logic it will be
    competitive and the solver that does not support it will be disadvantaged.
  }

  \item {\bf \mvaltrack.}  \emergencystretch.5cm Last year's competition introduced the experimental
  divisions QF\_Equality (only with QF\_UF benchmarks), QF\_Equality+LinearArith (only with
  QF\_UFIDL, QF\_UFLIA, and QF\_UFLRA benchmarks), and QF\_Equality+Bitvec (only with
  QF\_UFBV benchmarks).
  %
  This year these divisions will no longer be experimental.
  %
  In addition, we add this year the new experimental division QF\_FPArith
  (excluding the logics QF\_ABVFP, QF\_AUFBVFP, and QF\_ABVFPLRA).

  \rationale{Last year's divisions ran successfully, so there
    is no reason to keep them experimental.
    %
    For the new divisions, as before, given the inconsistencies across different
    model producing solvers, we proceed in an experimental fashion to push for
    model standardization. We exclude logics containing arrays since their model
    production adds further complications which we only intend to tackle in the
    future.}

  \item {\bf \prooftrack.}  \todo{after we iron it out.}

@@

\subsection{\prooftrack}

\todo{}

@@

\header{Disclosure.}
%
Haniel Barbosa is part of the developing teams of the SMT solvers
cvc5~\cite{cvc4-smtcomp18} \hb{update reference} and veriT~\cite{verit}.
%
Fran\c{c}ois is part of the developing team of the SMT solver COLIBRI~\cite{colibri}.
%
Jochen Hoenicke is part of the developing team of the SMT solver
SMTInterpol~\cite{smtinterpol}.
